Analyzing Toronto Island Ferry Ticket Counts

### 2024-09-26 outputs/paper/paper.qmd
added citations of used libraries and changed colors for the plots

library(tidyverse)
library(opendatatoronto)
library(here)

@@

Timestamp: The date and time when each ticket count was recorded.
Redemption.Count: The number of ferry tickets redeemed during the specified time period.
Sales.Count: The number of ferry tickets sold during the specified time period.

The data was processed and analyzed using the R programming language (@citeR). The tidyverse (@rohan) and opendatatoronto packages were employed to download, clean, and prepare the dataset for analysis. The resulting cleaned dataset was then utilized to generate the visualizations and statistical analyses presented in this paper.

## Results {#sec-data-results}

@@

## Sample Cleaned Data 
The table @tbl-cleaned_data provides a detailed snapshot of the Toronto Island Ferry ticket transactions for a specific time period. It includes a unique ID for each transaction, the Time at which each transaction occurred, the number of Tickets Redeemed, the number of Tickets Sold, and the Date of the transaction.
```{r}
#| label: tbl-cleaned_data
#| fig-cap: Sample Cleaned Data
#| echo: false

@@

The dataset used in this analysis includes timestamped records of ticket redemptions and sales for the Toronto Island Ferry system, offering a time series perspective on ferry usage. The plot in @fig-linePlot specifically highlights how these counts fluctuate over time, reflecting the flow of passenger demand. The y-axis for redemption counts is displayed in blue, while the y-axis for sales counts is displayed in green, but inverted to better show their correlation.
\newpage

@@

# Create the plot
ggplot(data, aes(x = Timestamp)) +
  # Line plot for Redemption Count
  geom_line(aes(y = Redemption.Count), color = "blue", linewidth = 1) +
  # Line plot for Sales Count inverted
  geom_line(aes(y = -Sales.Count), color = "green", linewidth = 1) +
  # Set y-axis limits symmetrically
  scale_y_continuous(limits = c(-y_max, y_max),
                     breaks = seq(-y_max, y_max, by = 1000),

                     sec.axis = sec_axis(~ -., 
                                         name = "Sales Count",
                                         breaks = seq(y_max, -y_max, by = -1000))) +
  labs(title = "Redemption and Sales Count Over Time",
       y = "Redemption Count",
       x = "Timestamp") +
  theme_minimal() +
  theme(axis.title.y.right = element_text(color = "green"),
        axis.title.y.left = element_text(color = "blue"))
```

@@

ggplot(data, aes(x=Sales.Count, y=Redemption.Count)) +
  geom_point(alpha=0.7) +
  labs(title="Scatterplot: Redemption Count vs Sales Count",
       x="Sales Count",
       y="Redemption Count") +
  xlim(0, max(data$Sales.Count) * 1.1) +  # Adding some space to the x-axis
  ylim(0, max(data$Redemption.Count) * 1.1) +  # Adding some space to the y-axis
  theme_minimal()

@@

ggplot(data, aes(x = "Sales Count", y = Sales.Count, fill = "Sales Count")) +
  geom_boxplot(alpha = 0.5) +
  geom_boxplot(aes(x = "Redemption Count", y = Redemption.Count, fill = "Redemption Count"), alpha = 0.5) +
  scale_fill_manual(values = c("Sales Count" = "lightblue", "Redemption Count" = "lightgreen")) +
  labs(title = "Box Plot of Sales and Redemption Counts",
       x = "Category",
       y = "Count",
       fill = "Legend") +
  theme_minimal() +
  theme(legend.position = "top")
```
@fig-boxPlot illustrates two box plots side by side: one for sales counts (light blue) and the other for redemption counts (light green). Each box plot provides a visual summary of the data distribution, highlighting the median, quartiles, and any outliers for each category.

>>>>

library(tidyverse)
library(opendatatoronto)
library(here)
library(ggplot2)

@@

**Timestamp**: The date and time when each ticket count was recorded.  
**Redemption.Count**: The number of ferry tickets redeemed during the specified time period.  
**Sales.Count**: The number of ferry tickets sold during the specified time period.

The data was processed and analyzed using the R programming language (@citeR). The tidyverse (@wickham2019tidyverse) and opendatatoronto packages were employed to download, clean, and prepare the dataset for analysis. The resulting cleaned dataset was then utilized to generate the visualizations and statistical analyses presented in this paper using a collection of libraries which included the following packages:
- ggplot2(@wickham2016ggplot2)
- knitr(@xie2023knitr)
- readr(@wickham2023readr)

## Results {#sec-data-resultsk}

@@

The table @tbl-cleaned_data provides a detailed snapshot of the Toronto Island Ferry ticket transactions for a specific time period. It includes a unique ID for each transaction, the Time at which each transaction occurred, the number of Tickets Redeemed, the number of Tickets Sold, and the Date of the transaction.
```{r}
#| label: tbl-cleaned_data
#| fig-cap: Sample Cleaned Data
#| echo: false

@@

The dataset used in this analysis includes timestamped records of ticket redemptions and sales for the Toronto Island Ferry system, offering a time series perspective on ferry usage. The plot in @fig-linePlot specifically highlights how these counts fluctuate over time, reflecting the flow of passenger demand. The y-axis for redemption counts is displayed in blue, while the y-axis for sales counts is displayed in green, but inverted to better show their correlation.

@@

# Create the plot
ggplot(data, aes(x = Timestamp)) +
  # Line plot for Redemption Count
  geom_line(aes(y = Redemption.Count), color = "#F7C4C1", linewidth = 1.2) +
  # Line plot for Sales Count (inverted)
  geom_line(aes(y = -Sales.Count), color = "#81B0B2", linewidth = 1.2) +

  # Set y-axis limits symmetrically and apply the secondary axis
  scale_y_continuous(limits = c(-y_max, y_max),
                     breaks = seq(-y_max, y_max, by = 1000),
                     sec.axis = sec_axis(~ -., 
                                         name = "Sales Count",
                                         breaks = seq(y_max, -y_max, by = -1000))) +

  # Add plot labels
  labs(
    title = "Redemption and Sales Count Over Time",
    y = "Redemption Count",  # Y-axis for Redemption Count
    x = "Timestamp"          # X-axis for Timestamp
  ) +

  # Apply minimal theme and customize axis titles
  theme_minimal() +
  theme(
    # Color the right-side axis title (Sales Count) to match the orange line
    axis.title.y.right = element_text(color = "#F7C4C1", size = 12),
    # Color the left-side axis title (Redemption Count) to match the pink line
    axis.title.y.left = element_text(color = "#81B0B2", size = 12),
    # Center the title and increase font size
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  )
```

@@

ggplot(data, aes(x = Sales.Count, y = Redemption.Count)) +
  geom_point(alpha = 0.7, color = "#93445e") +
  labs(
    title = "Scatterplot: Redemption Count vs Sales Count",
    x = "Sales Count",
    y = "Redemption Count"
  ) +
  xlim(0, max(data$Sales.Count) * 1.1) +  # Adding some space to the x-axis
  ylim(0, max(data$Redemption.Count) * 1.1) +  # Adding some space to the y-axis
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5)  # Center the title
  )
```

@@

ggplot(data, aes(x = "Sales Count", y = Sales.Count, fill = "Sales Count")) +
  geom_boxplot(alpha = 0.5) +
  geom_boxplot(aes(x = "Redemption Count", y = Redemption.Count, fill = "Redemption Count"), alpha = 0.5) +
  scale_fill_manual(values = c("Sales Count" = "lightgreen", "Redemption Count" = "lightblue")) +
  labs(title = "Box Plot of Sales and Redemption Counts",
       x = "Category",
       y = "Count",
       fill = "Legend") +
  theme_minimal() +
  theme(
    legend.position = "top",
    plot.title = element_text(hjust = 0.5)  # Center the title
  )
```
@fig-boxPlot illustrates two box plots side by side: one for sales counts (light blue) and the other for redemption counts (light green). Each box plot provides a visual summary of the data distribution, highlighting the median, quartiles, and any outliers for each category.
